---
title: "test"
author: "Nick McManus"
date: "2023-07-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(terra)
library(tidyverse)
library(lubridate)
library(here)
library(leaflet)
library(raster)
library(sf)
```


We want to find and save the zip codes that intersect with Kern county. We'll read in county and zip code vectors for the whole state, join them, then only select those zipcodes in Kern county. This .shp will then be exported for use in the Shiny app.
*Note*: originally I attempted this by using the `terra::intersect` tool, but this resulted in cutting off portions of zipcodes outside of Kern county. When trying to do the spatial analysis in the `sf` package, it keeps too many geometries that touch but don't intersect. Can work on this more in the future to determine best method.

```{r}
## Read in raw data
zips <- st_read(here('data/zipcodes/CA_Zips.shp'))
counties <- st_read(here('data/counties_ca/cnty19_1.shp')) 

## make sure both layers have same crs
zips_trans <- zips %>% 
  st_transform(zips, crs = st_crs(counties)) %>%
  ## only need to keep zipcode number
  dplyr::select(GEOID10) 

## Only keep name of counties
counties <- dplyr::select(counties, COUNTY_NAM)

# kern <- dplyr::filter(counties, COUNTY_NAM == "Kern") 


## Keep zipcodes that touch Kern county
kern_zips <- st_join(zips_trans, counties, join = st_intersects) %>% 
  dplyr::filter(COUNTY_NAM == "Kern", 
                ## remove select zips that barely touch Kern
                !GEOID10 %in% c(93536, 93535, 93201, 
                                93239, 93204, 93453,
                                93257, 93260, 93219, 
                                93261, 93218)) 
  
## save
st_write(kern_zips, here("data/zipcodes/kern_zips.shp"))

```

```{r}
### Method using combination of sf and terra:

# zips <- st_read(here('data/zipcodes/CA_Zips.shp'))
# counties <- st_read(here('data/counties_ca/cnty19_1.shp')) 
# 
# ## make sure both layers have same crs
# zips_trans <- zips %>% 
#   st_transform(zips, crs = st_crs(counties)) %>%
#   ## only need to keep zipcode number
#   dplyr::select(GEOID10) %>% 
#   vect()
# 
# ## isolate Kern county
# kern <- counties %>% 
#   dplyr::filter(COUNTY_NAM == "Kern") %>% 
#   vect()
# 
# kern_zips <- terra::intersect(kern, zips_trans)
# 
# crs(kern_zips) <- crs(trans_r)
# 
# writeVector(kern_zips, here('data/zipcodes/kern_zips.shp'), overwrite = T)
```



Now we'll determine how much standing water is present within each Kern county zip code. This data will be exported as a .csv and used for the Shiny app.
*Note:* Some standing water is outside of a zip code...

```{r}
## Read in standing water data at 30m resolution
water_jun <- rast(here('data/water/LC09_CU_003011_20230625_20230701_02_DSWE_Binary2_T.tif')) 
water_jun <- subst(water_jun, NA, 0)

## easier to transform vector than high-res raster
kern_zips <- vect(here("data/zipcodes/kern_zips.shp")) %>% 
  project(y = crs(water_jun))

## find the number of cells with water in each zipcode
water_zonal <- terra::zonal(water_jun, kern_zips, fun = 'sum', na.rm = T)

## add zipcode info to each zone
water_zips <- water_zonal %>% 
  mutate(zipcode = kern_zips$GEOID10) %>% 
  ## clean up names
  rename(water_ha = LC09_CU_003011_20230625_20230701_02_DSWE_Binary2_T) %>% 
  ## convert values from "units" ha
  ## w/30m resolution, each unit is 900 m^2
  ## 10,000 m^2 are in one hectare
  mutate(water_ha = water_ha * 900 / 10000) %>% 
  ## finally add date information
  mutate(date = ymd("2023-06-25"))

## export as .csv
write_csv(water_zips, here('data/water/water_ha_zipcode.csv'))
```












Testing leaflet outside of Shiny app

```{r}
trans_r <- raster(here('data/Kern_transmission_raster_wgs84.tif'))
pal <- colorNumeric(c("#0C2C84", "#41B6C4", "#FFFFCC"), values(trans_r),
                    na.color = "transparent")


leaflet() %>% 
  addRasterImage(trans_r, colors = pal, project = FALSE)

```

